# Does the antibiotic administration deplete the gut microbiota?
```{r data_antibiotics, comment="", message=FALSE, warning=FALSE, echo=FALSE}
load("data/data_16092025.Rdata")
load("data/objects_16092025.Rdata")
load("data/alpha_16092025.Rdata")
load("data/beta_16092025.Rdata")
```

## DNA extraction

```{r data_dna_extraction, comment="", message=FALSE, warning=FALSE, echo=FALSE}
dna_extraction <- read_excel("data/dna_extraction.xlsx")

dna_extraction_df<-dna_extraction %>%
  #pivot_longer(-sample, names_to = "metric", values_to = "value") %>%
  left_join(., sample_metadata,by=join_by(sample==Tube_code))
```

```{r data_dna_extraction_model, comment="", message=FALSE, warning=FALSE, echo=FALSE}
model_dna<- lmerTest::lmer(dna_extraction ~ time_point.x+(1|individual), data = dna_extraction_df)
summary(model_dna)
```

## Bacterial fraction change

```{r data_bacterial_fraction, comment="", message=FALSE, warning=FALSE, echo=FALSE}
mag_fragtion_accli_antib<-data_stats_filter %>%
  mutate(
    unmapped_reads = trimmed_reads - mapped_lizard - mapped_mags,
    mag_proportion = mapped_mags / (mapped_mags + unmapped_reads),
    singlem_read_fraction = singlem_read_fraction
  ) %>%
  select(sample, mag_proportion, singlem_read_fraction) %>%
  mutate(
    mag_proportion = if_else(singlem_read_fraction == 0, 0, mag_proportion),
    singlem_read_fraction = if_else(singlem_read_fraction == 0, NA, singlem_read_fraction),
    singlem_read_fraction = if_else(singlem_read_fraction < mag_proportion, NA, singlem_read_fraction),
    singlem_read_fraction = if_else(singlem_read_fraction > 1, 1, singlem_read_fraction)
  ) %>%
  pivot_longer(-sample, names_to = "proportion", values_to = "value") %>%
  mutate(
    proportion = factor(
      proportion,
      levels = c("mag_proportion", "singlem_read_fraction")
    )
  ) %>%
  left_join(., sample_metadata, by = join_by(sample == Tube_code)) %>%
  filter(!is.na(time_point))%>%
  filter(time_point %in% c("Acclimation", "Antibiotics") & proportion=="mag_proportion")
```

```{r data_bacterial_fraction_model, comment="", message=FALSE, warning=FALSE, echo=FALSE}
model_fraction<- lmerTest::lmer(value ~ time_point+(1|individual), data = mag_fragtion_accli_antib)
summary(model_fraction)
```


# Does the antibiotic administration change the microbial community?

## Alpha diversity
```{r filter_data, comment="", message=FALSE, warning=FALSE}
alpha_div_meta <- alpha_div %>%
  left_join(sample_metadata, by = join_by(sample == sample))%>%
  filter(time_point %in% c("Acclimation", "Antibiotics") ) 
```
```{r alpha_div_plot_1, comment="", message=FALSE, warning=FALSE, fig.height=12, fig.width=12, fig.fullwidth=TRUE}
label_map <- c(
  "richness" = "Species Richness",
  "neutral" = "Neutral Diversity",
  "phylogenetic" = "Phylogenetic Diversity")

alpha_div%>%
  pivot_longer(-sample, names_to = "metric", values_to = "value") %>%
  left_join(., sample_metadata, by = "sample") %>%
  mutate(metric = factor(metric, levels = c("richness", "neutral", "phylogenetic")))%>%
  #time_point = factor(time_point, levels = c("Wild", "Acclimation"))) %>%
  filter(time_point %in% c("Acclimation", "Antibiotics")) %>% 
  ggplot(aes(y = value, x = time_point, color=species, fill=species)) +
  geom_boxplot(width = 0.5,alpha=0.5, outlier.shape = NA, show.legend = FALSE) +
  geom_jitter(width = 0.2,show.legend = FALSE) +
  scale_color_manual(values=c('#008080', "#d57d2c")) +
  scale_fill_manual(values=c('#008080', "#d57d2c")) +
  facet_grid(metric ~ species, scales = "free_y", labeller = labeller(metric = label_map, type = label_map))+
  theme_classic() +
  theme(
    strip.background = element_blank(),
    panel.grid.minor.x = element_line(size = .1, color = "grey"),
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 18, lineheight = 0.6),
      ) +
  ylab("Alpha diversity")
```

***Richness***
```{r LMM, comment="", message=FALSE, warning=FALSE}
Modelq0GLMMNB <- glmer.nb(richness ~ time_point*species+(1|individual), data = alpha_div_meta)
summary(Modelq0GLMMNB)
```

***Neutral***
```{r LMM_neutral, comment="", message=FALSE, warning=FALSE}
Model_neutral <- nlme::lme(fixed = neutral ~ time_point*species, data = alpha_div_meta,
               random = ~ 1 | individual)
summary(Model_neutral)
```

***Phylogenetic***
```{r LMM_phylo, comment="", message=FALSE, warning=FALSE}
Model_phylo <- nlme::lme(fixed = phylogenetic ~ time_point*species, data = alpha_div_meta,
               random = ~ 1 | individual)
summary(Model_phylo)
```

## Beta diversity
```{r filter_data1, comment="", message=FALSE, warning=FALSE}
samples_to_keep <- sample_metadata %>%
  filter(time_point %in% c("Acclimation", "Antibiotics")) %>% 
  select(Tube_code) %>% 
  pull()
subset_meta <- sample_metadata %>%
  filter(time_point %in% c("Acclimation", "Antibiotics"))
```

***Richness***
```{r permanova_richness, comment="", message=FALSE, warning=FALSE}
richness <- as.matrix(beta_q0n$S)
richness <- as.dist(richness[rownames(richness) %in% samples_to_keep,
               colnames(richness) %in% samples_to_keep])

betadisper(richness, subset_meta$time_point) %>% permutest(.) 
```

```{r permanova_richness_adonis, comment="", message=FALSE, warning=FALSE}
adonis2(richness ~ time_point*species,
        data = subset_meta %>% arrange(match(Tube_code,labels(richness))),
        permutations = 999,
        strata = subset_meta %>% arrange(match(Tube_code,labels(richness))) %>% pull(individual),
        by="terms") %>%
        broom::tidy() %>%
        tt()
```

***Neutral***
```{r permanova_neutral, comment="", message=FALSE, warning=FALSE}
neutral <- as.matrix(beta_q1n$S)
neutral <- as.dist(neutral[rownames(neutral) %in% samples_to_keep,
               colnames(neutral) %in% samples_to_keep])

betadisper(neutral, subset_meta$time_point) %>% permutest(.) 
```

```{r permanova_neutral_adonis, comment="", message=FALSE, warning=FALSE}
adonis2(neutral ~ time_point*species,
        data = subset_meta %>% arrange(match(Tube_code,labels(neutral))),
        permutations = 999,
        strata = subset_meta %>% arrange(match(Tube_code,labels(neutral))) %>% pull(individual),
        by="terms") %>%
        broom::tidy() %>%
        tt()
```

***Phylogenetic***
```{r permanova_phylo, comment="", message=FALSE, warning=FALSE}
phylo <- as.matrix(beta_q1p$S)
phylo <- as.dist(phylo[rownames(phylo) %in% samples_to_keep,
               colnames(phylo) %in% samples_to_keep])
betadisper(phylo, subset_meta$time_point) %>% permutest(.)
```

```{r permanova_phylo_adonis, comment="", message=FALSE, warning=FALSE}
adonis2(phylo ~ time_point*species,
        data = subset_meta %>% arrange(match(Tube_code,labels(phylo))),
        permutations = 999,
        strata = subset_meta %>% arrange(match(Tube_code,labels(phylo))) %>% pull(individual),
        by="terms") %>%
        broom::tidy() %>%
        tt()
```

***dbRDA***
```{r rda_accli_antib, comment="", message=FALSE, warning=FALSE}

#Richness
cca_ord <- capscale(formula = richness ~ subset_meta$time_point* subset_meta$Population)
CAP_df <- as.data.frame(vegan::scores(cca_ord, display = "sites")) %>%
  rownames_to_column('Tube_code') %>%
  left_join(subset_meta, by = 'Tube_code') %>%
  column_to_rownames('Tube_code')%>%
  mutate(x_cen = mean(CAP1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(CAP2, na.rm = TRUE))

biplot_scores <- as.data.frame(vegan::scores(cca_ord, display = "bp")) %>%
  rownames_to_column("Variable")
biplot_scores$Variable <- recode(
  biplot_scores$Variable,
  "subset_meta$time_pointAntibiotics" = "Antibiotics",
  "subset_meta$PopulationWarm" = "Warm population",
  "subset_meta$time_pointAntibiotics:subset_meta$PopulationWarm" =
    "Interaction Wam population"
)

CAP_df %>%
  group_by(Population, time_point) %>%
  mutate(x_cen = mean(CAP1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(CAP2, na.rm = TRUE)) %>%
  ungroup() %>%
  ggplot(., aes(x=CAP1,y=CAP2, color=Population,shape = time_point)) +
  scale_color_manual(values=c('#008080', "#d57d2c")) +
  scale_fill_manual(values=c('#00808050', "#d57d2c50")) +
  geom_point(size=2) +
  geom_hline(yintercept = 0, linetype = "dotted") + 
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_segment(aes(x=x_cen, y=y_cen, xend=CAP1, yend=CAP2), alpha=0.2) +
  geom_segment(data = biplot_scores, aes(x = 0, y = 0, xend = CAP1, yend = CAP2),
               inherit.aes = FALSE, arrow = arrow(length = unit(0.2, "cm")), color = "black") +
  geom_text(data = biplot_scores, aes(x = CAP1, y = CAP2, label = Variable),
            inherit.aes = FALSE, color = "black", vjust = -0.5, hjust = 0.5)+
  theme_classic()

#Neutral
cca_ord <- capscale(formula = neutral ~ subset_meta$time_point* subset_meta$Population)
CAP_df <- as.data.frame(vegan::scores(cca_ord, display = "sites")) %>%
  rownames_to_column('Tube_code') %>%
  left_join(subset_meta, by = 'Tube_code') %>%
  column_to_rownames('Tube_code')%>%
  mutate(x_cen = mean(CAP1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(CAP2, na.rm = TRUE))

biplot_scores <- as.data.frame(vegan::scores(cca_ord, display = "bp")) %>%
  rownames_to_column("Variable")
biplot_scores$Variable <- recode(
  biplot_scores$Variable,
  "subset_meta$time_pointAntibiotics" = "Antibiotics",
  "subset_meta$PopulationWarm" = "Warm population",
  "subset_meta$time_pointAntibiotics:subset_meta$PopulationWarm" =
    "Interaction Wam population"
)

CAP_df %>%
  group_by(Population, time_point) %>%
  mutate(x_cen = mean(CAP1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(CAP2, na.rm = TRUE)) %>%
  ungroup() %>%
  ggplot(., aes(x=CAP1,y=CAP2, color=Population,shape = time_point)) +
  scale_color_manual(values=c('#008080', "#d57d2c")) +
    scale_fill_manual(values=c('#00808050', "#d57d2c50"))+
  geom_point(size=2) +
  geom_hline(yintercept = 0, linetype = "dotted") + 
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_segment(aes(x=x_cen, y=y_cen, xend=CAP1, yend=CAP2), alpha=0.2) +
  geom_segment(data = biplot_scores, aes(x = 0, y = 0, xend = CAP1, yend = CAP2),
               inherit.aes = FALSE, arrow = arrow(length = unit(0.2, "cm")), color = "black") +
  geom_text(data = biplot_scores, aes(x = CAP1, y = CAP2, label = Variable),
            inherit.aes = FALSE, color = "black", vjust = -0.5, hjust = 0.5)+
  theme_classic()


#Phylogenetic
cca_ord <- capscale(formula = phylo ~ subset_meta$time_point* subset_meta$Population)
CAP_df <- as.data.frame(vegan::scores(cca_ord, display = "sites")) %>%
  rownames_to_column('Tube_code') %>%
  left_join(subset_meta, by = 'Tube_code') %>%
  column_to_rownames('Tube_code')%>%
  mutate(x_cen = mean(CAP1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(CAP2, na.rm = TRUE))

biplot_scores <- as.data.frame(vegan::scores(cca_ord, display = "bp")) %>%
  rownames_to_column("Variable")
biplot_scores$Variable <- recode(
  biplot_scores$Variable,
  "subset_meta$time_pointAntibiotics" = "Antibiotics",
  "subset_meta$PopulationWarm" = "Warm population",
  "subset_meta$time_pointAntibiotics:subset_meta$PopulationWarm" =
    "Interaction Wam population"
)

CAP_df %>%
  group_by(Population, time_point) %>%
  mutate(x_cen = mean(CAP1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(CAP2, na.rm = TRUE)) %>%
  ungroup() %>%
  ggplot(., aes(x=CAP1,y=CAP2, color=Population,shape = time_point)) +
  scale_color_manual(values=c('#008080', "#d57d2c")) +
  scale_fill_manual(values=c('#00808050', "#d57d2c50")) +
  geom_point(size=2) +
  geom_hline(yintercept = 0, linetype = "dotted") + 
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_segment(aes(x=x_cen, y=y_cen, xend=CAP1, yend=CAP2), alpha=0.2) +
  geom_segment(data = biplot_scores, aes(x = 0, y = 0, xend = CAP1, yend = CAP2),
               inherit.aes = FALSE, arrow = arrow(length = unit(0.2, "cm")), color = "black") +
  geom_text(data = biplot_scores, aes(x = CAP1, y = CAP2, label = Variable),
            inherit.aes = FALSE, color = "black", vjust = -0.5, hjust = 0.5)+
  theme_classic()
```

# Does antibiotic administration remove the differences found in the two populations?

## Alpha diversity

***Shapiro test***
```{r shapiro_antib_1, comment="", message=FALSE, warning=FALSE}
alpha_div %>%
  pivot_longer(-sample, names_to = "metric", values_to = "value") %>%
  left_join(., sample_metadata, by = join_by(sample == Tube_code)) %>%
  filter(time_point == "Antibiotics" ) %>%
  filter(metric=="richness") %>% 
  summarize(shapiro_p_value = shapiro.test(value)$p.value) %>%
  pull(shapiro_p_value)

alpha_div %>%
  pivot_longer(-sample, names_to = "metric", values_to = "value") %>%
  left_join(., sample_metadata, by = join_by(sample == Tube_code)) %>%
  filter(time_point == "Antibiotics" ) %>%
  filter(metric=="neutral") %>% 
  summarize(shapiro_p_value = shapiro.test(value)$p.value) %>%
  pull(shapiro_p_value)

alpha_div %>%
  pivot_longer(-sample, names_to = "metric", values_to = "value") %>%
  left_join(., sample_metadata, by = join_by(sample == Tube_code)) %>%
  filter(time_point == "Antibiotics" ) %>%
  filter(metric=="phylogenetic") %>% 
  summarize(shapiro_p_value = shapiro.test(value)$p.value) %>%
  pull(shapiro_p_value)
```

```{r filter_data01, comment="", message=FALSE, warning=FALSE}
alpha_div_meta <- alpha_div %>%
  left_join(sample_metadata, by = join_by(sample == sample))%>%
  filter(time_point == "Antibiotics" )
```

```{r alpha_div_plot_antib, comment="", message=FALSE, warning=FALSE}
alpha_div %>%
  pivot_longer(-sample, names_to = "metric", values_to = "value") %>%
  left_join(., sample_metadata, by = "sample") %>%
  mutate(metric=factor(metric,levels=c("richness", "neutral"))) %>%
  filter(metric!="phylogenetic") %>%
  filter(time_point == "Antibiotics" ) %>% 
  ggplot(aes(y = value, x = species, color=species, fill=species)) +
  geom_boxplot(width = 0.5,alpha=0.5 ,outlier.shape = NA, show.legend = FALSE) +
  geom_jitter(width = 0.2, show.legend = FALSE) +
  scale_color_manual(values=c('#008080', "#d57d2c")) +
  scale_fill_manual(values=c('#008080', "#d57d2c")) +
  stat_compare_means(method = "t.test", show.legend = F, size = 3, label.x = c(1.5))+
  facet_grid( ~ metric, scales = "free_y")+
  theme_classic() +
  theme(
    strip.background = element_blank(),
    panel.grid.minor.x = element_line(size = .1, color = "grey"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 16, lineheight = 0.6)
      ) +
  ylab("Alpha diversity")
```

```{r alpha_div_plot_2, comment="", message=FALSE, warning=FALSE}
alpha_div %>%
  pivot_longer(-sample, names_to = "metric", values_to = "value") %>%
  left_join(., sample_metadata, by = "sample") %>%
  filter(metric=="phylogenetic") %>%
  filter(time_point == "Antibiotics" ) %>% 
   ggplot(aes(y = value, x = species, color=species, fill=species)) +
  geom_boxplot(width = 0.5,alpha=0.5 ,outlier.shape = NA, show.legend = FALSE) +
  geom_jitter(width = 0.2, show.legend = FALSE) +
  scale_color_manual(values=c('#008080', "#d57d2c")) +
      scale_fill_manual(values=c('#008080', "#d57d2c")) +
  stat_compare_means(method = "wilcox.test", show.legend = F, size = 3, label.x = c(1.5))+
  facet_grid( ~ metric, scales = "free_y")+
      theme_classic() +
        theme(
    strip.background = element_blank(),
    panel.grid.minor.x = element_line(size = .1, color = "grey"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 16, lineheight = 0.6)
      ) +
  ylab("Alpha diversity")
```

## Beta diversity
```{r data_filter1, comment="", message=FALSE, warning=FALSE}
samples_to_keep <- sample_metadata %>%
  filter(time_point == "Antibiotics") %>% 
  select(Tube_code) %>% 
  pull()
subset_meta <- sample_metadata %>%
  filter(time_point == "Antibiotics")
```

***Richness***
```{r permanova_rich1, comment="", message=FALSE, warning=FALSE}
richness <- as.matrix(beta_q0n$S)
richness <- as.dist(richness[rownames(richness) %in% samples_to_keep,
               colnames(richness) %in% samples_to_keep])

betadisper(richness, subset_meta$species) %>% permutest(.) 
```

```{r permanova_rich1_adonis, comment="", message=FALSE, warning=FALSE}
adonis2(richness ~ species,
        data = subset_meta %>% arrange(match(Tube_code,labels(richness))),
        permutations = 999) %>%
        broom::tidy() %>%
        tt()
```

***Neutral***
```{r permanova_neutral1, comment="", message=FALSE, warning=FALSE}
neutral <- as.matrix(beta_q1n$S)
neutral <- as.dist(neutral[rownames(neutral) %in% samples_to_keep,
               colnames(neutral) %in% samples_to_keep])

betadisper(neutral, subset_meta$species) %>% permutest(.)
```

```{r permanova_neutral1_adonis, comment="", message=FALSE, warning=FALSE}
adonis2(neutral ~ species,
        data = subset_meta %>% arrange(match(Tube_code,labels(neutral))),
        permutations = 999) %>%
        broom::tidy() %>%
        tt()
```

***Phylogenetic***
```{r permanova_phylo1, comment="", message=FALSE, warning=FALSE}
phylo <- as.matrix(beta_q1p$S)
phylo <- as.dist(phylo[rownames(phylo) %in% samples_to_keep,
               colnames(phylo) %in% samples_to_keep])
betadisper(phylo, subset_meta$species) %>% permutest(.) 
```
```{r permanova_phylo1_adonis, comment="", message=FALSE, warning=FALSE}
adonis2(phylo ~ species,
        data = subset_meta %>% arrange(match(Tube_code,labels(phylo))),
        permutations = 999) %>%
        broom::tidy() %>%
        tt()
```

```{r betad_div_plot, comment="", message=FALSE, warning=FALSE}
richness %>%
  metaMDS(.,trymax = 500, k=2, verbosity=FALSE, trace=FALSE) %>%
  vegan::scores() %>%
  as_tibble(., rownames = "sample") %>%
  left_join(subset_meta, by = join_by(sample == Tube_code)) %>%
  group_by(species, time_point) %>%
  mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
  ungroup() %>%
  ggplot(., aes(x=NMDS1,y=NMDS2, color=species)) +
  scale_color_manual(values=c('#008080',"#d57d2c")) +
  geom_point(size = 2) +
  geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
  labs(y = "NMDS2", x="NMDS1 \n Richness beta diversity") +
  theme_classic() +
  theme()

neutral %>%
  metaMDS(.,trymax = 500, k=2, verbosity=FALSE, trace=FALSE) %>%
  vegan::scores() %>%
  as_tibble(., rownames = "sample") %>%
  left_join(subset_meta, by = join_by(sample == Tube_code))%>%
  group_by(species, time_point) %>%
  mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
  ungroup() %>%
  ggplot(., aes(x=NMDS1,y=NMDS2, color=species)) +
  scale_color_manual(values=c('#008080',"#d57d2c")) +
  geom_point(size = 2) +
  geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
  labs(y = "NMDS2", x = "NMDS1 \n Neutral beta diversity") +
  theme_classic() +
  theme()

phylo %>%
  metaMDS(.,trymax = 500, k=2, verbosity=FALSE, trace=FALSE) %>%
  vegan::scores() %>%
  as_tibble(., rownames = "sample") %>%
  left_join(subset_meta, by = join_by(sample == Tube_code))%>%
  group_by(species, time_point) %>%
  mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
  ungroup() %>%
  ggplot(., aes(x=NMDS1,y=NMDS2, color=species)) +
  scale_color_manual(values = c('#008080', "#d57d2c"))+
  geom_point(size=2) +
  geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
  labs(y= element_blank (),x="NMDS1 \n Phylogenetic beta diversity") +
  theme_classic() +
  theme() 
```
