# Faecal Microbiota Transplantation (FMT)

## What is the trend of the microbiota in each type after FMT?

### Alpha diversity

```{r alpha_div_plot_all, comment="", message=FALSE, warning=FALSE, fig.height=10, fig.width=8, fig.fullwidth=TRUE}
label_map <- c(
  "Control" = "Cold-control", 
  "Treatment" = "Cold-intervention",
  "Hot_control" = "Warm-control",
  "richness" = "Species Richness",
  "neutral" = "Neutral Diversity",
  "phylogenetic" = "Phylogenetic Diversity"
)

alpha_div %>%
  pivot_longer(-sample, names_to = "metric", values_to = "value") %>%
  left_join(., sample_metadata, by = join_by(sample == Tube_code)) %>%
  filter(time_point %in% c("FMT1","Acclimation", "FMT2")) %>%
  mutate(metric = factor(metric, levels = c("richness", "neutral", "phylogenetic"))) %>%
  ggplot(aes(y = value, x = time_point, color = type, fill = type)) +
  geom_boxplot(outlier.shape = NA, show.legend = FALSE) +
  geom_point(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8), alpha = 0.5, show.legend = FALSE) +
  facet_grid(metric ~ type, scales = "free_y", labeller = labeller(metric = label_map, type = label_map))+
#  facet_nested(.~metric+type, labeller = labeller(metric = label_map, type = label_map))+
  scale_color_manual(name="Type",
                       breaks=c("Cold_control", "Warm_control", "Cold_intervention"),
                       labels=c("Cold_control", "Warm_control", "Cold_intervention"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
      scale_fill_manual(name="Type",
                       breaks=c("Cold_control", "Warm_control", "Cold_intervention"),
                       labels=c("Cold_control", "Warm_control", "Cold_intervention"),
                       values=c("#4477AA50","#d57d2c50","#76b18350")) +
  theme_minimal() +
  theme(axis.text.x=element_text(size=10))+
  labs(x = "Time Point", y = "value")
```

```{r 2}
alpha_div_meta <- alpha_div %>%
  left_join(sample_metadata, by = join_by(sample == Tube_code)) %>%
  filter(time_point=="FMT1"|time_point=="Acclimation"|time_point=="FMT2") 
```

***Richness***
```{r alpha_div_data_post7_comparison, comment="", message=FALSE, warning=FALSE}
Modelq0GLMMNB <- glmer.nb(richness ~ type*time_point+(1|individual), data = alpha_div_meta)
summary(Modelq0GLMMNB)
emmeans::emmeans(Modelq0GLMMNB, pairwise ~ type)
emmeans::emmeans(Modelq0GLMMNB, pairwise ~ time_point)
```

***Neutral***
```{r alpha_div_data_1_post7_comparison, comment="", message=FALSE, warning=FALSE}
Modelq1n <- nlme::lme(neutral ~ type*time_point, 
               random = ~ 1 | individual,
               data = alpha_div_meta)
summary(Modelq1n)
emmeans::emmeans(Modelq1n, pairwise ~ type)
emmeans::emmeans(Modelq1n, pairwise ~ time_point)
```

***Phylogenetic***
```{r alpha_div_data_2_post7_comparison, comment="", message=FALSE, warning=FALSE}
Model_phylo <- nlme::lme(phylogenetic ~ type*time_point, 
               random = ~ 1 | individual,
               data = alpha_div_meta)
summary(Model_phylo)
emmeans::emmeans(Model_phylo, pairwise ~ type)
emmeans::emmeans(Model_phylo, pairwise ~ time_point)
```

#### Beta diversity

***Number of samples used***
```{r beta_div_neutral_all_samples_2_comparison, results = TRUE, message=FALSE, warning=FALSE, comments=FALSE}
samples_to_keep_post7 <- sample_metadata %>%
  filter(time_point=="FMT1"|time_point=="Acclimation"|time_point=="FMT2") %>% 
  select(Tube_code) %>% 
  pull()
subset_meta_post7 <- sample_metadata %>%
  filter(time_point=="FMT1"|time_point=="Acclimation"|time_point=="FMT2")

subset_meta_post7$time_point<-as.factor(subset_meta_post7$time_point)
subset_meta_post7$type<-as.factor(subset_meta_post7$type)
length(samples_to_keep_post7)
```

***Richness***
```{r beta_richness_permanova_post7_2_comparison, warning=FALSE, comments="", message=FALSE}
richness_post7 <- as.matrix(beta_q0n$S)
richness_post7 <- as.dist(richness_post7[rownames(richness_post7) %in% samples_to_keep_post7,
               colnames(richness_post7) %in% samples_to_keep_post7])
betadisper(richness_post7, subset_meta_post7$type) %>% permutest(., pairwise = TRUE)
```

```{r beta_richness_permanova_post7_1_comparison, warning=FALSE, comments="", message=FALSE}
adonis2(richness_post7 ~ type*time_point,
        data = subset_meta_post7 %>% arrange(match(Tube_code,labels(richness_post7))),
        permutations = 999,
        strata = subset_meta_post7 %>% arrange(match(Tube_code,labels(richness_post7))) %>% pull(individual),
        by="terms") %>%
        broom::tidy() %>%
        tt()
```

***Neutral***
```{r beta_neutral_permanova_post7_2_comparison, warning=FALSE, comments="", message=FALSE}
neutral_post7 <- as.matrix(beta_q1n$S)
neutral_post7 <- as.dist(neutral_post7[rownames(neutral_post7) %in% samples_to_keep_post7,
               colnames(neutral_post7) %in% samples_to_keep_post7])
betadisper(neutral_post7, subset_meta_post7$type) %>% permutest(., pairwise = TRUE)
```

```{r beta_neutral_permanova_post7_1_comparison, warning=FALSE, comments="", message=FALSE}
adonis2(neutral_post7 ~ type*time_point,
        data = subset_meta_post7 %>% arrange(match(Tube_code,labels(neutral_post7))),
        permutations = 999,
        strata = subset_meta_post7 %>% arrange(match(Tube_code,labels(neutral_post7))) %>% pull(individual),
        by="terms") %>%        
        broom::tidy() %>%
        tt()
```

***Phylogenetic***
```{r beta_phylo_permanova_post7_2_comparison, warning=FALSE, comments="", message=FALSE}
phylo_post7 <- as.matrix(beta_q1p$S)
phylo_post7 <- as.dist(phylo_post7[rownames(phylo_post7) %in% samples_to_keep_post7,
               colnames(phylo_post7) %in% samples_to_keep_post7])
betadisper(phylo_post7, subset_meta_post7$type) %>% permutest(., pairwise = TRUE)
```

```{r beta_phylo_permanova_post7_1_comparison, warning=FALSE, comments="", message=FALSE}
adonis2(phylo_post7 ~ type*time_point,
        data = subset_meta_post7 %>% arrange(match(Tube_code,labels(phylo_post7))),
        permutations = 999,
        strata = subset_meta_post7 %>% arrange(match(Tube_code,labels(phylo_post7))) %>% pull(individual),
          by="terms") %>%        
        broom::tidy() %>%
        tt()
```

***dbRDA***
```{r beta_neutral_nmds_post7_comparison, warning=FALSE, comments="", message=FALSE, results="hide"}
#Richness
cca_ord <- capscale(formula = richness_post7 ~ subset_meta_post7$time_point* subset_meta_post7$type)
CAP_df <- as.data.frame(vegan::scores(cca_ord, display = "sites")) %>%
  rownames_to_column('Tube_code') %>%
  left_join(subset_meta_post7, by = 'Tube_code') %>%
  column_to_rownames('Tube_code')%>%
  mutate(x_cen = mean(CAP1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(CAP2, na.rm = TRUE))

biplot_scores <- as.data.frame(vegan::scores(cca_ord, display = "bp")) %>%
  rownames_to_column("Variable")
biplot_scores$Variable <- recode(biplot_scores$Variable, 
                                 "subset_meta_post7$time_pointFMT1" = "FMT1",
                                 "subset_meta_post7$time_pointFMT2"="FMT2",
                                 "subset_meta_post7$typeHot_control"="Warm-control",
                                 "subset_meta_post7$typeTreatment" = "Cold-intervention",
                                 "subset_meta_post7$time_pointFMT1:subset_meta_post7$typeHot_control"="FMT1 Warm-control",
                                 "subset_meta_post7$time_pointFMT2:subset_meta_post7$typeHot_control"="FMT2 Warm-control",
                                 "subset_meta_post7$time_pointFMT1:subset_meta_post7$typeTreatment" = "FMT1 Cold-intervention",
                                 "subset_meta_post7$time_pointFMT2:subset_meta_post7$typeTreatment" = "FMT2 Cold-intervention")

beta_richness_nmds_post7 <-CAP_df %>%
  group_by(type, time_point) %>%
  mutate(x_cen = mean(CAP1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(CAP2, na.rm = TRUE)) %>%
  ungroup() %>%
  ggplot(., aes(x=CAP1,y=CAP2, color=type,shape = time_point)) +
  scale_color_manual(name="Type",
          breaks=c("Cold_control", "Warm_control", "Cold_intervention"),
          labels=c("Cold-control","Warm-control", "Cold-intervention"),
          values=c("#4477AA","#d57d2c", "#76b183")) +
      scale_fill_manual(name="Type",
          breaks=c("Cold_control", "Warm_control", "Cold_intervention"),
          labels=c("Cold-control","Warm-control", "Cold-intervention"),
          values=c("#4477AA50","#d57d2c50","#76b18350")) +
        scale_x_discrete(labels = c("Control" = "Cold-Cold", "Hot_control" = "Hot-Hot", "Treatment" = "Cold-Hot")) +
  geom_point(size=2) +
  geom_hline(yintercept = 0, linetype = "dotted") + 
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_segment(aes(x=x_cen, y=y_cen, xend=CAP1, yend=CAP2), alpha=0.2) +
  geom_segment(data = biplot_scores, aes(x = 0, y = 0, xend = CAP1, yend = CAP2),
               inherit.aes = FALSE, arrow = arrow(length = unit(0.2, "cm")), color = "black") +
  geom_text(data = biplot_scores, aes(x = CAP1, y = CAP2, label = Variable),
            inherit.aes = FALSE, color = "black", vjust = -0.5, hjust = 0.5)+
  theme_classic()

#Neutral
cca_ord <- capscale(formula = neutral_post7 ~ subset_meta_post7$time_point* subset_meta_post7$type)
CAP_df <- as.data.frame(vegan::scores(cca_ord, display = "sites")) %>%
  rownames_to_column('Tube_code') %>%
  left_join(subset_meta_post7, by = 'Tube_code') %>%
  column_to_rownames('Tube_code')%>%
  mutate(x_cen = mean(CAP1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(CAP2, na.rm = TRUE))

biplot_scores <- as.data.frame(vegan::scores(cca_ord, display = "bp")) %>%
  rownames_to_column("Variable")
biplot_scores$Variable <- recode(biplot_scores$Variable, 
                                 "subset_meta_post7$time_pointFMT1" = "FMT1",
                                 "subset_meta_post7$time_pointFMT2"="FMT2",
                                 "subset_meta_post7$typeHot_control"="Warm-control",
                                 "subset_meta_post7$typeTreatment" = "Cold-intervention",
                                 "subset_meta_post7$time_pointFMT1:subset_meta_post7$typeHot_control"="FMT1 Warm-control",
                                 "subset_meta_post7$time_pointFMT2:subset_meta_post7$typeHot_control"="FMT2 Warm-control",
                                 "subset_meta_post7$time_pointFMT1:subset_meta_post7$typeTreatment" = "FMT1 Cold-intervention",
                                 "subset_meta_post7$time_pointFMT2:subset_meta_post7$typeTreatment" = "FMT2 Cold-intervention")

beta_neutral_nmds_post7 <- CAP_df %>%
  group_by(type, time_point) %>%
  mutate(x_cen = mean(CAP1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(CAP2, na.rm = TRUE)) %>%
  ungroup() %>%
  ggplot(., aes(x=CAP1,y=CAP2, color=type,shape = time_point)) +
  scale_color_manual(name="Type",
          breaks=c("Cold_control", "Warm_control", "Cold_intervention"),
          labels=c("Cold-control","Warm-control", "Cold-intervention"),
          values=c("#4477AA","#d57d2c", "#76b183")) +
      scale_fill_manual(name="Type",
          breaks=c("Cold_control", "Warm_control", "Cold_intervention"),
          labels=c("Cold-control","Warm-control", "Cold-intervention"),
          values=c("#4477AA50","#d57d2c50","#76b18350")) +
        scale_x_discrete(labels = c("Control" = "Cold-Cold", "Hot_control" = "Hot-Hot", "Treatment" = "Cold-Hot")) +
  geom_point(size=2) +
  geom_hline(yintercept = 0, linetype = "dotted") + 
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_segment(aes(x=x_cen, y=y_cen, xend=CAP1, yend=CAP2), alpha=0.2) +
  geom_segment(data = biplot_scores, aes(x = 0, y = 0, xend = CAP1, yend = CAP2),
               inherit.aes = FALSE, arrow = arrow(length = unit(0.2, "cm")), color = "black") +
  geom_text(data = biplot_scores, aes(x = CAP1, y = CAP2, label = Variable),
            inherit.aes = FALSE, color = "black", vjust = -0.5, hjust = 0.5)+
  theme_classic()


#Phylogenetic
cca_ord <- capscale(formula = phylo_post7 ~ subset_meta_post7$time_point* subset_meta_post7$type)
CAP_df <- as.data.frame(vegan::scores(cca_ord, display = "sites")) %>%
  rownames_to_column('Tube_code') %>%
  left_join(subset_meta_post7, by = 'Tube_code') %>%
  column_to_rownames('Tube_code')%>%
  mutate(x_cen = mean(CAP1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(CAP2, na.rm = TRUE))

biplot_scores <- as.data.frame(vegan::scores(cca_ord, display = "bp")) %>%
  rownames_to_column("Variable")
biplot_scores$Variable <- recode(biplot_scores$Variable, 
                                 "subset_meta_post7$time_pointFMT1" = "FMT1",
                                 "subset_meta_post7$time_pointFMT2"="FMT2",
                                 "subset_meta_post7$typeHot_control"="Warm-control",
                                 "subset_meta_post7$typeTreatment" = "Cold-intervention",
                                 "subset_meta_post7$time_pointFMT1:subset_meta_post7$typeHot_control"="FMT1 Warm-control",
                                 "subset_meta_post7$time_pointFMT2:subset_meta_post7$typeHot_control"="FMT2 Warm-control",
                                 "subset_meta_post7$time_pointFMT1:subset_meta_post7$typeTreatment" = "FMT1 Cold-intervention",
                                 "subset_meta_post7$time_pointFMT2:subset_meta_post7$typeTreatment" = "FMT2 Cold-intervention")

beta_phylogenetic_nmds_post7 <- CAP_df %>%
  group_by(type, time_point) %>%
  mutate(x_cen = mean(CAP1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(CAP2, na.rm = TRUE)) %>%
  ungroup() %>%
  ggplot(., aes(x=CAP1,y=CAP2, color=type,shape = time_point)) +
  scale_color_manual(name="Type",
          breaks=c("Cold_control", "Warm_control", "Cold_intervention"),
          labels=c("Cold-control","Warm-control", "Cold-intervention"),
          values=c("#4477AA","#d57d2c", "#76b183")) +
      scale_fill_manual(name="Type",
          breaks=c("Cold_control", "Warm_control", "Cold_intervention"),
          labels=c("Cold-control","Warm-control", "Cold-intervention"),
          values=c("#4477AA50","#d57d2c50","#76b18350")) +
        scale_x_discrete(labels = c("Control" = "Cold-Cold", "Hot_control" = "Hot-Hot", "Treatment" = "Cold-Hot")) +
  geom_point(size=2) +
  geom_hline(yintercept = 0, linetype = "dotted") + 
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_segment(aes(x=x_cen, y=y_cen, xend=CAP1, yend=CAP2), alpha=0.2) +
  geom_segment(data = biplot_scores, aes(x = 0, y = 0, xend = CAP1, yend = CAP2),
               inherit.aes = FALSE, arrow = arrow(length = unit(0.2, "cm")), color = "black") +
  geom_text(data = biplot_scores, aes(x = CAP1, y = CAP2, label = Variable),
            inherit.aes = FALSE, color = "black", vjust = -0.5, hjust = 0.5)+
  theme_classic()
```

```{r beta_neutral_nmds_plot_post7_final, fig.height=8, fig.width=12, fig.fullwidth=TRUE}
ggarrange(beta_richness_nmds_post7, beta_neutral_nmds_post7, beta_phylogenetic_nmds_post7, ncol=3, nrow=1, common.legend = TRUE, legend="right")
```
```{r beta_neutral_nmds_plot_post7_final_1, fig.height=8, fig.width=12, fig.fullwidth=TRUE}
beta_neutral_nmds_post7
```

```{r data_FMT_1, comment="", message=FALSE, warning=FALSE, echo=FALSE}
load("data/data_03122025.Rdata")
load("data/objects_16092025.Rdata")
load("data/alpha_16092025.Rdata")
load("data/beta_16092025.Rdata")
```

## Differences between acclimation and FMT1 across the three experimental groups

### Beta diversity

***CI from acclimation to FMT1***
```{r beta_div_accli_fmt1_ci, comment="", message=FALSE, warning=FALSE}

samples_to_keep <- sample_metadata %>%
  filter(time_point %in% c("FMT1","Acclimation") & type == "Cold_intervention") %>%
  select(Tube_code) %>% 
  pull()
subset_meta <- sample_metadata %>%
  filter(time_point %in% c("FMT1","Acclimation") & type == "Cold_intervention")

length(samples_to_keep)
```

```{r beta_div_accli_fmt1_ci_1, comment="", message=FALSE, warning=FALSE}

richness <- as.matrix(beta_q0n$S)
richness <- as.dist(richness[rownames(richness) %in% samples_to_keep,
               colnames(richness) %in% samples_to_keep])
betadisper(richness, subset_meta$time_point) %>% permutest(., pairwise = TRUE)
```

```{r beta_div_accli_fmt1_ci_6, comment="", message=FALSE, warning=FALSE}
adonis2(richness ~ time_point,
        data = subset_meta %>% arrange(match(Tube_code,labels(richness))),
        permutations = 999,
        strata = subset_meta %>% arrange(match(Tube_code,labels(richness))) %>% pull(individual),
        by="terms") %>%
        broom::tidy() %>%
        tt()
```        

```{r beta_div_accli_fmt1_ci_2, comment="", message=FALSE, warning=FALSE}

neutral <- as.matrix(beta_q1n$S)
neutral <- as.dist(neutral[rownames(neutral) %in% samples_to_keep,
               colnames(neutral) %in% samples_to_keep])
betadisper(neutral, subset_meta$time_point) %>% permutest(., pairwise = TRUE)
```    

```{r beta_div_accli_fmt1_ci_5, comment="", message=FALSE, warning=FALSE}
adonis2(neutral ~ time_point,
        data = subset_meta %>% arrange(match(Tube_code,labels(neutral))),
        strata = subset_meta %>% arrange(match(Tube_code,labels(neutral))) %>% pull(individual),
        by="terms") %>%
        broom::tidy() %>%
        tt()
```   

```{r beta_div_accli_fmt1_ci_3, comment="", message=FALSE, warning=FALSE}
phylo <- as.matrix(beta_q1p$S)
phylo <- as.dist(phylo[rownames(phylo) %in% samples_to_keep,
               colnames(phylo) %in% samples_to_keep])
betadisper(phylo, subset_meta$time_point) %>% permutest(., pairwise = TRUE)

```   
```{r beta_div_accli_fmt1_ci_4, comment="", message=FALSE, warning=FALSE}

adonis2(phylo ~ time_point,
        data = subset_meta %>% arrange(match(Tube_code,labels(phylo))),
        permutations = 999,
        strata = subset_meta %>% arrange(match(Tube_code,labels(phylo))) %>% pull(individual),
        by="terms") %>%
        tt()
```


***CC from acclimation to FMT1***
```{r beta_div_accli_fmt1_cc, comment="", message=FALSE, warning=FALSE}

samples_to_keep <- sample_metadata %>%
  filter(time_point %in% c("FMT1","Acclimation") & type == "Cold_control") %>%
  select(Tube_code) %>% 
  pull()
subset_meta <- sample_metadata %>%
  filter(time_point %in% c("FMT1","Acclimation") & type == "Cold_control")

length(samples_to_keep)
```

```{r beta_div_accli_fmt1_cc_1, comment="", message=FALSE, warning=FALSE}

richness <- as.matrix(beta_q0n$S)
richness <- as.dist(richness[rownames(richness) %in% samples_to_keep,
               colnames(richness) %in% samples_to_keep])
betadisper(richness, subset_meta$time_point) %>% permutest(., pairwise = TRUE)
```

```{r beta_div_accli_fmt1_cc_2, comment="", message=FALSE, warning=FALSE}
adonis2(richness ~ time_point,
        data = subset_meta %>% arrange(match(Tube_code,labels(richness))),
        permutations = 999,
        strata = subset_meta %>% arrange(match(Tube_code,labels(richness))) %>% pull(individual),
        by="terms") %>%
        broom::tidy() %>%
        tt()
```        

```{r beta_div_accli_fmt1_cc_3, comment="", message=FALSE, warning=FALSE}

neutral <- as.matrix(beta_q1n$S)
neutral <- as.dist(neutral[rownames(neutral) %in% samples_to_keep,
               colnames(neutral) %in% samples_to_keep])
betadisper(neutral, subset_meta$time_point) %>% permutest(., pairwise = TRUE)
```    

```{r beta_div_accli_fmt1_cc_4, comment="", message=FALSE, warning=FALSE}
adonis2(neutral ~ time_point,
        data = subset_meta %>% arrange(match(Tube_code,labels(neutral))),
        strata = subset_meta %>% arrange(match(Tube_code,labels(neutral))) %>% pull(individual),
        by="terms") %>%
        broom::tidy() %>%
        tt()
```   

```{r beta_div_accli_fmt1_cc_5, comment="", message=FALSE, warning=FALSE}
phylo <- as.matrix(beta_q1p$S)
phylo <- as.dist(phylo[rownames(phylo) %in% samples_to_keep,
               colnames(phylo) %in% samples_to_keep])
betadisper(phylo, subset_meta$time_point) %>% permutest(., pairwise = TRUE)

```   
```{r beta_div_accli_fmt1_cc_6, comment="", message=FALSE, warning=FALSE}

adonis2(phylo ~ time_point,
        data = subset_meta %>% arrange(match(Tube_code,labels(phylo))),
        permutations = 999,
        strata = subset_meta %>% arrange(match(Tube_code,labels(phylo))) %>% pull(individual),
        by="terms") %>%
        tt()
```

***WC from acclimation to FMT1***
```{r beta_div_accli_fmt1_wc, comment="", message=FALSE, warning=FALSE}

samples_to_keep <- sample_metadata %>%
  filter(time_point %in% c("FMT1","Acclimation") & type == "Warm_control") %>%
  select(Tube_code) %>% 
  pull()
subset_meta <- sample_metadata %>%
  filter(time_point %in% c("FMT1","Acclimation") & type == "Warm_control")

length(samples_to_keep)
```

```{r beta_div_accli_fmt1_wc_1, comment="", message=FALSE, warning=FALSE}

richness <- as.matrix(beta_q0n$S)
richness <- as.dist(richness[rownames(richness) %in% samples_to_keep,
               colnames(richness) %in% samples_to_keep])
betadisper(richness, subset_meta$time_point) %>% permutest(., pairwise = TRUE)
```

```{r beta_div_accli_fmt1_wc_2, comment="", message=FALSE, warning=FALSE}
adonis2(richness ~ time_point,
        data = subset_meta %>% arrange(match(Tube_code,labels(richness))),
        permutations = 999,
        strata = subset_meta %>% arrange(match(Tube_code,labels(richness))) %>% pull(individual),
        by="terms") %>%
        broom::tidy() %>%
        tt()
```        

```{r beta_div_accli_fmt1_wc_3, comment="", message=FALSE, warning=FALSE}

neutral <- as.matrix(beta_q1n$S)
neutral <- as.dist(neutral[rownames(neutral) %in% samples_to_keep,
               colnames(neutral) %in% samples_to_keep])
betadisper(neutral, subset_meta$time_point) %>% permutest(., pairwise = TRUE)
```    

```{r beta_div_accli_fmt1_wc_4, comment="", message=FALSE, warning=FALSE}
adonis2(neutral ~ time_point,
        data = subset_meta %>% arrange(match(Tube_code,labels(neutral))),
        strata = subset_meta %>% arrange(match(Tube_code,labels(neutral))) %>% pull(individual),
        by="terms") %>%
        broom::tidy() %>%
        tt()
```   

```{r beta_div_accli_fmt1_wc_5, comment="", message=FALSE, warning=FALSE}
phylo <- as.matrix(beta_q1p$S)
phylo <- as.dist(phylo[rownames(phylo) %in% samples_to_keep,
               colnames(phylo) %in% samples_to_keep])
betadisper(phylo, subset_meta$time_point) %>% permutest(., pairwise = TRUE)

```   
```{r beta_div_accli_fmt1_wc_6, comment="", message=FALSE, warning=FALSE}

adonis2(phylo ~ time_point,
        data = subset_meta %>% arrange(match(Tube_code,labels(phylo))),
        permutations = 999,
        strata = subset_meta %>% arrange(match(Tube_code,labels(phylo))) %>% pull(individual),
        by="terms") %>%
        tt()
```



### Functional differences

***CI from acclimation to FMT1***

```{r comunity_elem_FMT3_cold_1, comment="", message=FALSE, warning=FALSE}
significant_element<-element_gift %>%
  filter(time_point %in% c("FMT1","Acclimation") & type == "Cold_intervention") %>%
  select(-time_point, -type) %>%
  filter(rowSums(across(where(is.numeric), ~ . != 0)) > 0) %>%
  select(Tube_code, where( ~ is.numeric(.) && sum(.) > 0)) %>%
  left_join(sample_metadata[c(1, 8)], by = "Tube_code") %>% 
  pivot_longer(-c(Tube_code, time_point),
               names_to = "trait",
               values_to = "value") %>%
  group_by(trait) %>%
  summarise(p_value = wilcox.test(value ~ time_point)$p.value) %>%
  mutate(p_adjust = p.adjust(p_value, method = "BH")) %>%
  filter(p_adjust < 0.05) %>%
  remove_rownames() %>% 
  left_join(., uniqueGIFT_db, by=join_by("trait"=="Code_element"))
```

```{r comunity_elem_fm2_3, comment="", message=FALSE, warning=FALSE}
element_gift_sig <- element_gift %>%
  select(Tube_code, all_of(intersect(
    significant_element$trait,
    colnames(element_gift)
  ))) %>%
  left_join(., sample_metadata[c(1, 6, 8 )], by = join_by(Tube_code == Tube_code)) %>%
  filter(time_point %in% c("FMT1","Acclimation"))%>%
  filter(type=="Cold_intervention")
  
difference_table <- element_gift_sig %>%
  select(-Tube_code, -type) %>%
  group_by(time_point) %>%
  summarise(across(everything(), mean)) %>%
  t() %>%
  row_to_names(row_number = 1) %>%
  as.data.frame() %>%
  mutate_if(is.character, as.numeric) %>%
  rownames_to_column(., "Elements") %>%
  left_join(.,uniqueGIFT_db[c(1,3,4)],by = join_by(Elements == Code_element)) %>% 
  arrange(Function) %>% 
  mutate(Difference=Acclimation-FMT1)%>% 
  mutate(group_color = ifelse(Difference <0,"FMT1", "Acclimation")) 
```

```{r commun_wilcox_elem_plot_fmt2_1, comment="", message=FALSE, warning=FALSE, fig.height=6, fig.width=12, fig.fullwidth=TRUE}
difference_table %>%
  ggplot(aes(x=forcats::fct_reorder(Function,Difference), y=Difference, fill=group_color)) + 
  geom_col() +
  scale_fill_manual(values=c("#76b183",'#008080')) + 
  geom_hline(yintercept=0) + 
  coord_flip()+
  theme(axis.text = element_text(size = 10),
        axis.title = element_text(size = 12),
        legend.position = "right", 
        legend.title = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_line(size = 0.15, linetype = 'solid',
                                        colour = "grey"))+
  xlab("Function") + 
  ylab("Mean difference")
```

***CC from acclimation to FMT1***

```{r comunity_elem_FMT3_cold, comment="", message=FALSE, warning=FALSE}
significant_element<-element_gift %>%
  filter(time_point %in% c("FMT1","Acclimation") & type == "Cold_control") %>%
  select(-time_point, -type) %>%
  filter(rowSums(across(where(is.numeric), ~ . != 0)) > 0) %>%
  select(Tube_code, where( ~ is.numeric(.) && sum(.) > 0)) %>%
  left_join(sample_metadata[c(1, 8)], by = "Tube_code") %>% 
  pivot_longer(-c(Tube_code, time_point),
               names_to = "trait",
               values_to = "value") %>%
  group_by(trait) %>%
  summarise(p_value = wilcox.test(value ~ time_point)$p.value) %>%
  mutate(p_adjust = p.adjust(p_value, method = "BH")) %>%
  filter(p_adjust < 0.05) %>%
  remove_rownames() %>% 
  left_join(., uniqueGIFT_db, by=join_by("trait"=="Code_element"))
```

```{r comunity_elem_fm2_5, comment="", message=FALSE, warning=FALSE}
element_gift_sig <- element_gift %>%
  select(Tube_code, all_of(intersect(
    significant_element$trait,
    colnames(element_gift)
  ))) %>%
  left_join(., sample_metadata[c(1, 6, 8)], by = join_by(Tube_code == Tube_code)) %>%
  filter(time_point %in% c("FMT1","Acclimation"))%>%
  filter(type=="Cold_control")
  
difference_table <- element_gift_sig %>%
  select(-Tube_code, -type) %>%
  group_by(time_point) %>%
  summarise(across(everything(), mean)) %>%
  t() %>%
  row_to_names(row_number = 1) %>%
  as.data.frame() %>%
  mutate_if(is.character, as.numeric) %>%
  rownames_to_column(., "Elements") %>%
  left_join(.,uniqueGIFT_db[c(1,3,4)],by = join_by(Elements == Code_element)) %>% 
  arrange(Function) %>% 
  mutate(Difference=Acclimation-FMT1)%>% 
  mutate(group_color = ifelse(Difference <0,"FMT1", "Acclimation")) 
```

```{r commun_wilcox_elem_plot_fmt2_5, comment="", message=FALSE, warning=FALSE, fig.height=6, fig.width=12, fig.fullwidth=TRUE}
difference_table %>%
  ggplot(aes(x=forcats::fct_reorder(Function,Difference), y=Difference, fill=group_color)) + 
  geom_col() +
  scale_fill_manual(values=c("#4477AA",'#008080')) + 
  geom_hline(yintercept=0) + 
  coord_flip()+
  theme(axis.text = element_text(size = 10),
        axis.title = element_text(size = 12),
        legend.position = "right", 
        legend.title = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_line(size = 0.15, linetype = 'solid',
                                        colour = "grey"))+
  xlab("Function") + 
  ylab("Mean difference")
```

***WC from acclimation to FMT1***

```{r comunity_elem_FMT3_cold_2, comment="", message=FALSE, warning=FALSE}
element_gift %>%
  filter(time_point %in% c("FMT1","Acclimation") & type == "Warm_control") %>%
  select(-time_point, -type) %>%
  filter(rowSums(across(where(is.numeric), ~ . != 0)) > 0) %>%
  select(Tube_code, where( ~ is.numeric(.) && sum(.) > 0)) %>%
  left_join(sample_metadata[c(1, 8)], by = "Tube_code") %>% 
  pivot_longer(-c(Tube_code, time_point),
               names_to = "trait",
               values_to = "value") %>%
  group_by(trait) %>%
  summarise(p_value = wilcox.test(value ~ time_point)$p.value) %>%
  mutate(p_adjust = p.adjust(p_value, method = "BH")) %>%
  filter(p_adjust < 0.05) %>%
  remove_rownames() %>% 
  left_join(., uniqueGIFT_db, by=join_by("trait"=="Code_element"))
```

## Differences between FMT1 and FMT2 across the three experimental groups

### Functional differences

***CI from FMT1 to FMT2***

```{r comunity_elem_FMT3_cold_6, comment="", message=FALSE, warning=FALSE}
significant_element<-element_gift %>%
  filter(time_point %in% c("FMT2","FMT1") & type == "Cold_intervention") %>%
  select(-time_point, -type) %>%
  filter(rowSums(across(where(is.numeric), ~ . != 0)) > 0) %>%
  select(Tube_code, where( ~ is.numeric(.) && sum(.) > 0)) %>%
  left_join(sample_metadata[c(1, 8)], by = "Tube_code") %>% 
  pivot_longer(-c(Tube_code, time_point),
               names_to = "trait",
               values_to = "value") %>%
  group_by(trait) %>%
  summarise(p_value = wilcox.test(value ~ time_point)$p.value) %>%
  mutate(p_adjust = p.adjust(p_value, method = "BH")) %>%
  filter(p_adjust < 0.05) %>%
  remove_rownames() %>% 
  left_join(., uniqueGIFT_db, by=join_by("trait"=="Code_element"))
```

```{r comunity_elem_fm2_2, comment="", message=FALSE, warning=FALSE}
element_gift_sig <- element_gift %>%
  select(Tube_code, all_of(intersect(
    significant_element$trait,
    colnames(element_gift)
  ))) %>%
  left_join(., sample_metadata[c(1, 6, 8)], by = join_by(Tube_code == Tube_code)) %>%
  filter(time_point %in% c("FMT2","FMT1"))%>%
  filter(type=="Cold_intervention")
  
difference_table <- element_gift_sig %>%
  select(-Tube_code, -type) %>%
  group_by(time_point) %>%
  summarise(across(everything(), mean)) %>%
  t() %>%
  row_to_names(row_number = 1) %>%
  as.data.frame() %>%
  mutate_if(is.character, as.numeric) %>%
  rownames_to_column(., "Elements") %>%
  left_join(.,uniqueGIFT_db[c(1,3,4)],by = join_by(Elements == Code_element)) %>% 
  arrange(Function) %>% 
  mutate(Difference=FMT1-FMT2)%>% 
  mutate(group_color = ifelse(Difference <0,"FMT2", "FMT1")) 
```

```{r commun_wilcox_elem_plot_fmt2_2, comment="", message=FALSE, warning=FALSE, fig.height=6, fig.width=12, fig.fullwidth=TRUE}
difference_table %>%
  ggplot(aes(x=forcats::fct_reorder(Function,Difference), y=Difference, fill=group_color)) + 
  geom_col() +
  scale_fill_manual(values=c('#008080',"#76b183")) + 
  geom_hline(yintercept=0) + 
  coord_flip()+
  theme(axis.text = element_text(size = 10),
        axis.title = element_text(size = 12),
        legend.position = "right", 
        legend.title = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_line(size = 0.15, linetype = 'solid',
                                        colour = "grey"))+
  xlab("Function") + 
  ylab("Mean difference")
```

***CC from FMT1 to FMT2***

```{r comunity_elem_FMT3_cold_5, comment="", message=FALSE, warning=FALSE}
element_gift %>%
  filter(time_point %in% c("FMT2","FMT1") & type == "Cold_control") %>%
  select(-time_point, -type) %>%
  filter(rowSums(across(where(is.numeric), ~ . != 0)) > 0) %>%
  select(Tube_code, where( ~ is.numeric(.) && sum(.) > 0)) %>%
  left_join(sample_metadata[c(1, 8)], by = "Tube_code") %>% 
  pivot_longer(-c(Tube_code, time_point),
               names_to = "trait",
               values_to = "value") %>%
  group_by(trait) %>%
  summarise(p_value = wilcox.test(value ~ time_point)$p.value) %>%
  mutate(p_adjust = p.adjust(p_value, method = "BH")) %>%
  filter(p_adjust < 0.05) %>%
  remove_rownames() %>% 
  left_join(., uniqueGIFT_db, by=join_by("trait"=="Code_element"))
```

***WC from FMT1 to FMT2***

```{r comunity_elem_FMT3_cold_post6, comment="", message=FALSE, warning=FALSE}
element_gift %>%
  filter(time_point %in% c("FMT2","FMT1") & type == "Warm_control") %>%
  select(-time_point, -type) %>%
  filter(rowSums(across(where(is.numeric), ~ . != 0)) > 0) %>%
  select(Tube_code, where( ~ is.numeric(.) && sum(.) > 0)) %>%
  left_join(sample_metadata[c(1, 8)], by = "Tube_code") %>% 
  pivot_longer(-c(Tube_code, time_point),
               names_to = "trait",
               values_to = "value") %>%
  group_by(trait) %>%
  summarise(p_value = wilcox.test(value ~ time_point)$p.value) %>%
  mutate(p_adjust = p.adjust(p_value, method = "BH")) %>%
  filter(p_adjust < 0.05) %>%
  remove_rownames() %>% 
  left_join(., uniqueGIFT_db, by=join_by("trait"=="Code_element"))
```

# What are the differences at FMT1 between groups?

## Is the GM of the intervention group more similar to warm population after one week from FMT?

### Shapiro test

```{r shapiro_fmt1_1, comment="", message=FALSE, warning=FALSE,}
alpha_div %>%
  pivot_longer(-sample, names_to = "metric", values_to = "value") %>%
  left_join(., sample_metadata, by = join_by(sample == Tube_code)) %>%
  filter(time_point =="FMT1") %>% 
  filter(metric=="richness") %>% 
  summarize(shapiro_p_value = shapiro.test(value)$p.value) %>%
  pull(shapiro_p_value)

alpha_div %>%
  pivot_longer(-sample, names_to = "metric", values_to = "value") %>%
  left_join(., sample_metadata, by = join_by(sample == Tube_code)) %>%
  filter(time_point =="FMT1") %>% 
  filter(metric=="neutral") %>% 
  summarize(shapiro_p_value = shapiro.test(value)$p.value) %>%
  pull(shapiro_p_value)

alpha_div %>%
  pivot_longer(-sample, names_to = "metric", values_to = "value") %>%
  left_join(., sample_metadata, by = join_by(sample == Tube_code)) %>%
  filter(time_point =="FMT1") %>% 
  filter(metric=="phylogenetic") %>% 
  summarize(shapiro_p_value = shapiro.test(value)$p.value) %>%
  pull(shapiro_p_value)
```

### Alpha diversity
```{r alpha_div_plot_5_1, comment="", message=FALSE, warning=FALSE, fig.height=4, fig.width=10, fig.fullwidth=TRUE}
alpha_div %>%
  pivot_longer(-sample, names_to = "metric", values_to = "value") %>%
  left_join(., sample_metadata, by = join_by(sample == Tube_code)) %>%
  filter(time_point=="FMT1") %>%
  mutate(metric=factor(metric,levels=c("richness","neutral","phylogenetic"))) %>%
      ggplot(aes(y = value, x = type, group=type, color=type, fill=type)) +
      geom_boxplot(outlier.shape = NA) +
      geom_jitter(alpha=0.5) +
      scale_color_manual(name="Type",
          breaks=c("Cold_control", "Warm_control", "Cold_intervention"),
                       labels=c("Cold_control", "Warm_control", "Cold_intervention"),
          values=c("#4477AA","#d57d2c", "#76b183")) +
      scale_fill_manual(name="Type",
          breaks=c("Cold_control", "Warm_control", "Cold_intervention"),
                       labels=c("Cold_control", "Warm_control", "Cold_intervention"),
          values=c("#4477AA50","#d57d2c50","#76b18350")) +
      facet_wrap(. ~ metric,scales = "free") +
      coord_cartesian(xlim = c(1, NA)) +
      stat_compare_means(method="t.test",size=3, label.x=.7) +
      theme_classic() +
        theme(
    strip.background = element_blank(),
    panel.grid.minor.x = element_line(size = .1, color = "grey"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size=10),
    axis.text.x = element_blank(),
    # Increase plot size
    plot.title = element_text(size = 10),
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 8)
      ) +
  ylab("Alpha diversity")
```
***p<0.05 when comparing neutral beta diversity between Cold-control and Warm-control

### Beta diversity

***Number of samples used***
```{r beta_div_neutral_treat1_samples_2, results = TRUE, message=FALSE, warning=FALSE, comments=FALSE}
samples_to_keep_post1 <- sample_metadata %>%
  filter(time_point == "FMT1") %>% 
  select(Tube_code) %>% 
  pull()
subset_meta_post1 <- sample_metadata %>%
  filter(time_point == "FMT1")
length(samples_to_keep_post1)
```

***Richness***
```{r beta_richness_permanova_post1_2, warning=FALSE, comments="", message=FALSE}
richness_post1 <- as.matrix(beta_q0n$S)
richness_post1 <- as.dist(richness_post1[rownames(richness_post1) %in% samples_to_keep_post1,
               colnames(richness_post1) %in% samples_to_keep_post1])
betadisper(richness_post1, subset_meta_post1$type) %>% permutest(., pairwise = TRUE)
```

```{r beta_richness_permanova_post1_1, warning=FALSE, comments="", message=FALSE}
adonis2(richness_post1 ~ type,
        data = subset_meta_post1 %>% arrange(match(Tube_code,labels(richness_post1))),
        permutations = 999) %>%
        broom::tidy() %>%
        tt()
```

```{r beta_richness_permanova_post1_3, warning=FALSE, comments="", message=FALSE}
subset_meta_post1_arrange <- column_to_rownames(subset_meta_post1, "Tube_code")
subset_meta_post1_arrange<-subset_meta_post1_arrange[labels(richness_post1),]

pairwise <- pairwise.adonis(richness_post1, subset_meta_post1_arrange$type, perm=999)
pairwise%>%
  tt()
```

***Neutral***
```{r beta_neutral_permanova_post1_2, warning=FALSE, comments="", message=FALSE}
neutral_post1 <- as.matrix(beta_q1n$S)
neutral_post1 <- as.dist(neutral_post1[rownames(neutral_post1) %in% samples_to_keep_post1,
               colnames(neutral_post1) %in% samples_to_keep_post1])
betadisper(neutral_post1, subset_meta_post1$type) %>% permutest(., pairwise = TRUE)
```

```{r beta_neutral_permanova_post1_1, warning=FALSE, comments="", message=FALSE}
adonis2(neutral_post1 ~ type,
        data = subset_meta_post1 %>% arrange(match(Tube_code,labels(neutral_post1))),
        permutations = 999) %>%
        broom::tidy() %>%
        tt()
```

```{r beta_neutral_permanova_post1_3, warning=FALSE, comments="", message=FALSE}
subset_meta_post1_arrange <- column_to_rownames(subset_meta_post1, "Tube_code")
subset_meta_post1_arrange<-subset_meta_post1_arrange[labels(neutral_post1),]
pairwise <- pairwise.adonis(neutral_post1, subset_meta_post1_arrange$type, perm=999)
pairwise%>%
  tt()
```

***Phylogenetic***
```{r beta_phylo_permanova_post1_2, warning=FALSE, comments="", message=FALSE}
phylo_post1 <- as.matrix(beta_q1p$S)
phylo_post1 <- as.dist(phylo_post1[rownames(phylo_post1) %in% samples_to_keep_post1,
               colnames(phylo_post1) %in% samples_to_keep_post1])
betadisper(phylo_post1, subset_meta_post1$type) %>% permutest(., pairwise = TRUE)
```

```{r beta_phylo_permanova_post1_1, warning=FALSE, comments="", message=FALSE}
adonis2(phylo_post1 ~ type,
        data = subset_meta_post1 %>% arrange(match(Tube_code,labels(phylo_post1))),
        permutations = 999) %>%
        broom::tidy() %>%
        tt()
```

```{r beta_phylopermanova_post1_3, warning=FALSE, comments="", message=FALSE}
subset_meta_post1_arrange <- column_to_rownames(subset_meta_post1, "Tube_code")
subset_meta_post1_arrange<-subset_meta_post1_arrange[labels(phylo_post1),]
pairwise <- pairwise.adonis(phylo_post1, subset_meta_post1_arrange$type, perm=999)
pairwise%>%
  tt()
```

***NMDS***
```{r beta_neutral_nmds_post1, warning=FALSE, comments="", message=FALSE, results="hide"}
richness_post1 %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE, trace = FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(subset_meta_post1, by = join_by(sample == Tube_code)) %>%
			group_by(type) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type)) +
        scale_color_manual(name="Type",
                       breaks=c("Cold_control", "Warm_control", "Cold_intervention"),
                       labels=c("Cold_control", "Warm_control", "Cold_intervention"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y = "NMDS2", x="NMDS1 \n Richness beta diversity") +
				theme_classic() +
				theme(legend.position="none")

neutral_post1 %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE, trace = FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(subset_meta_post1, by = join_by(sample == Tube_code))%>%
			group_by(type) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type)) +
        scale_color_manual(name="Type",
                       breaks=c("Cold_control", "Warm_control", "Cold_intervention"),
                       labels=c("Cold_control", "Warm_control", "Cold_intervention"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y = "NMDS2", x="NMDS1 \n Neutral beta diversity") +
				theme_classic() +
				theme(legend.position="none")

phylo_post1 %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE, trace = FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(subset_meta_post1, by = join_by(sample == Tube_code))%>%
			group_by(type) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type)) +
				scale_color_manual(name="Type",
                       breaks=c("Cold_control", "Warm_control", "Cold_intervention"),
                       labels=c("Cold_control", "Warm_control", "Cold_intervention"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y= element_blank (), x="NMDS1 \n Phylogenetic beta diversity") +
				theme_classic() +
				theme(legend.position="none")
```

#### Functional differences
***MCI***
```{r gitfs_elements_tm5, warning=FALSE, comments="", message=FALSE}
GIFTs_functions_community %>%
  rowMeans() %>%
  as_tibble(., rownames = "sample") %>%
  left_join(sample_metadata, by = join_by(sample == Tube_code)) %>%
  filter(time_point=="FMT1") %>%
  group_by(type) %>%
  summarise(MCI = mean(value), sd = sd(value))
```

```{r gitfs_functional_tm5_stats_2, warning=FALSE, comments="", message=FALSE}
MCI_tm5_ele <- GIFTs_functions_community %>%
  rowMeans() %>%
  as_tibble(., rownames = "sample") %>%
  left_join(sample_metadata, by = join_by(sample == Tube_code)) %>%
  filter(time_point == "FMT1")

shapiro.test(MCI_tm5_ele$value)#normality test

res.aov <- aov(value ~ type, data = MCI_tm5_ele)#anova test
summary(res.aov)
```

***Functional differences***

CC and CI
```{r comunity_elem_7_1, comment="", message=FALSE, warning=FALSE}
element_gift %>%
  filter(type %in% c("Cold_control", "Cold_intervention") & time_point == "FMT1") %>%
  select(-time_point, -type) %>%
  filter(rowSums(across(where(is.numeric), ~ . != 0)) > 0) %>%
  select(Tube_code, where( ~ is.numeric(.) && sum(.) > 0)) %>%
  left_join(sample_metadata[c(1, 6)], by = "Tube_code") %>% 
  pivot_longer(-c(Tube_code, type),
               names_to = "trait",
               values_to = "value") %>%
  group_by(trait) %>%
  summarise(p_value = wilcox.test(value ~ type)$p.value) %>%
  mutate(p_adjust = p.adjust(p_value, method = "BH")) %>%
  filter(p_adjust < 0.05) %>%
  remove_rownames()
```

CI and WC
```{r comunity_elem_CI_WC, comment="", message=FALSE, warning=FALSE}
element_gift %>%
  filter(type %in% c("Warm_control", "Cold_intervention") &
           time_point == "FMT1") %>%
  select(-time_point, -type) %>%
  filter(rowSums(across(where(is.numeric), ~ . != 0)) > 0) %>%
  select(Tube_code, where(~ is.numeric(.) && sum(.) > 0)) %>%
  left_join(sample_metadata[c(1, 6)], by = "Tube_code") %>% 
  pivot_longer(-c(Tube_code,type), names_to = "trait", values_to = "value") %>%
  group_by(trait) %>%
  summarise(p_value = wilcox.test(value ~ type)$p.value) %>%
  mutate(p_adjust=p.adjust(p_value, method="BH")) %>%
  filter(p_adjust < 0.05)%>%
  remove_rownames()
```

CC and WC
```{r comunity_elem_CC_WC, comment="", message=FALSE, warning=FALSE}
element_gift %>%
  filter(type %in% c("Warm_control", "Cold_control") &
           time_point == "FMT1") %>%
  select(-time_point, -type) %>%
  filter(rowSums(across(where(is.numeric), ~ . != 0)) > 0) %>%
  select(Tube_code, where(~ is.numeric(.) && sum(.) > 0)) %>%
  left_join(sample_metadata[c(1, 6)], by = "Tube_code") %>%
  pivot_longer(-c(Tube_code,type), names_to = "trait", values_to = "value") %>%
  group_by(trait) %>%
  summarise(p_value = wilcox.test(value ~ type)$p.value) %>%
  mutate(p_adjust=p.adjust(p_value, method="BH")) %>%
  filter(p_adjust < 0.05)%>%
  remove_rownames()
```

## Is the GM of the intervention group still more similar to warm population after two weeks from FMT?

### Shapiro test

```{r shapiro_fmt2, comment="", message=FALSE, warning=FALSE,}
alpha_div %>%
  pivot_longer(-sample, names_to = "metric", values_to = "value") %>%
  left_join(., sample_metadata, by = join_by(sample == Tube_code)) %>%
  filter(time_point =="FMT2") %>% 
  filter(metric=="richness") %>% 
  summarize(shapiro_p_value = shapiro.test(value)$p.value) %>%
  pull(shapiro_p_value)

alpha_div %>%
  pivot_longer(-sample, names_to = "metric", values_to = "value") %>%
  left_join(., sample_metadata, by = join_by(sample == Tube_code)) %>%
  filter(time_point =="FMT2") %>% 
  filter(metric=="neutral") %>% 
  summarize(shapiro_p_value = shapiro.test(value)$p.value) %>%
  pull(shapiro_p_value)

alpha_div %>%
  pivot_longer(-sample, names_to = "metric", values_to = "value") %>%
  left_join(., sample_metadata, by = join_by(sample == Tube_code)) %>%
  filter(time_point =="FMT2") %>% 
  filter(metric=="phylogenetic") %>% 
  summarize(shapiro_p_value = shapiro.test(value)$p.value) %>%
  pull(shapiro_p_value)
```

### Alpha diversity
```{r alpha_div_plot_FMT2, comment="", message=FALSE, warning=FALSE, fig.height=4, fig.width=10, fig.fullwidth=TRUE}
alpha_div %>%
  pivot_longer(-sample, names_to = "metric", values_to = "value") %>%
  left_join(., sample_metadata, by = join_by(sample == Tube_code)) %>%
  filter(time_point=="FMT2") %>%
  mutate(metric=factor(metric,levels=c("richness","neutral","phylogenetic"))) %>%
      ggplot(aes(y = value, x = type, group=type, color=type, fill=type)) +
      geom_boxplot(outlier.shape = NA) +
      geom_jitter(alpha=0.5) +
      scale_color_manual(name="Type",
          breaks=c("Cold_control", "Warm_control", "Cold_intervention"),
                       labels=c("Cold_control", "Warm_control", "Cold_intervention"),
          values=c("#4477AA","#d57d2c", "#76b183")) +
      scale_fill_manual(name="Type",
          breaks=c("Cold_control", "Warm_control", "Cold_intervention"),
                       labels=c("Cold_control", "Warm_control", "Cold_intervention"),
          values=c("#4477AA50","#d57d2c50","#76b18350")) +
        scale_x_discrete(labels = c("Control" = "Cold-Cold", "Hot_control" = "Hot-Hot", "Treatment" = "Cold-Hot")) +
      facet_wrap(. ~ metric,scales = "free") +
      coord_cartesian(xlim = c(1, NA)) +
      stat_compare_means(method="t.test",size=3, label.x=.7) +
      theme_classic() +
        theme(
    strip.background = element_blank(),
    panel.grid.minor.x = element_line(size = .1, color = "grey"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size=10),
    axis.text.x = element_blank(),
    # Increase plot size
    plot.title = element_text(size = 10),
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 8)
      ) +
  ylab("Alpha diversity")

```
***p<0.05 when comparing richness and neutral beta diversity between Cold-control and Cold-intervention


### Beta diversity

***Number of samples used***
```{r beta_div_neutral_FMT2, results = TRUE, message=FALSE, warning=FALSE, comments=FALSE}
samples_to_keep_FMT2 <- sample_metadata %>%
  filter(time_point == "FMT2") %>% 
  select(Tube_code) %>% 
  pull()
subset_meta_FMT2 <- sample_metadata %>%
  filter(time_point == "FMT2")
length(samples_to_keep_FMT2)
```

***Richness***
```{r beta_richness_permanova_FMT2_2, warning=FALSE, comments="", message=FALSE}
richness_FMT2 <- as.matrix(beta_q0n$S)
richness_FMT2 <- as.dist(richness_FMT2[rownames(richness_FMT2) %in% samples_to_keep_FMT2,
               colnames(richness_FMT2) %in% samples_to_keep_FMT2])
betadisper(richness_FMT2, subset_meta_FMT2$type) %>% permutest(., pairwise = TRUE)
```

```{r beta_richness_permanova_FMT2_1, warning=FALSE, comments="", message=FALSE}
adonis2(richness_FMT2 ~ type,
        data = subset_meta_FMT2 %>% arrange(match(Tube_code,labels(richness_FMT2))),
        permutations = 999) %>%
        broom::tidy() %>%
        tt()
```

```{r beta_richness_permanova_FMT2_3, warning=FALSE, comments="", message=FALSE}
subset_meta_FMT2_arrange <- column_to_rownames(subset_meta_FMT2, "Tube_code")
subset_meta_FMT2_arrange<-subset_meta_FMT2_arrange[labels(richness_FMT2),]

pairwise <- pairwise.adonis(richness_FMT2, subset_meta_FMT2_arrange$type, perm=999)
pairwise%>%
  tt()
```

***Neutral***
```{r beta_neutral_permanova_FMT2_2, warning=FALSE, comments="", message=FALSE}
neutral_FMT2 <- as.matrix(beta_q1n$S)
neutral_FMT2 <- as.dist(neutral_FMT2[rownames(neutral_FMT2) %in% samples_to_keep_FMT2,
               colnames(neutral_FMT2) %in% samples_to_keep_FMT2])
betadisper(neutral_FMT2, subset_meta_FMT2$type) %>% permutest(., pairwise = TRUE)
```

```{r beta_neutral_permanova_FMT2_1, warning=FALSE, comments="", message=FALSE}
adonis2(neutral_FMT2 ~ type,
        data = subset_meta_FMT2 %>% arrange(match(Tube_code,labels(neutral_FMT2))),
        permutations = 999) %>%
        broom::tidy() %>%
        tt()
```

```{r beta_neutral_permanova_FMT2_3, warning=FALSE, comments="", message=FALSE}
subset_meta_FMT2_arrange <- column_to_rownames(subset_meta_FMT2, "Tube_code")
subset_meta_FMT2_arrange<-subset_meta_FMT2_arrange[labels(neutral_FMT2),]
pairwise <- pairwise.adonis(neutral_FMT2, subset_meta_FMT2_arrange$type, perm=999)
pairwise%>%
  tt()
```

***Phylogenetic***
```{r beta_phylo_permanova_FMT2_2, warning=FALSE, comments="", message=FALSE}
phylo_FMT2 <- as.matrix(beta_q1p$S)
phylo_FMT2 <- as.dist(phylo_FMT2[rownames(phylo_FMT2) %in% samples_to_keep_FMT2,
               colnames(phylo_FMT2) %in% samples_to_keep_FMT2])
betadisper(phylo_FMT2, subset_meta_FMT2$type) %>% permutest(., pairwise = TRUE)
```

```{r beta_phylo_permanova_FMT2_1, warning=FALSE, comments="", message=FALSE}
adonis2(phylo_FMT2 ~ type,
        data = subset_meta_FMT2 %>% arrange(match(Tube_code,labels(phylo_FMT2))),
        permutations = 999) %>%
        broom::tidy() %>%
        tt()
```
```{r beta_phylo_permanova_FMT2_pair, warning=FALSE, comments="", message=FALSE}
subset_meta_FMT2_arrange <- column_to_rownames(subset_meta_FMT2, "Tube_code")
subset_meta_FMT2_arrange<-subset_meta_FMT2_arrange[labels(phylo_FMT2),]
pairwise <- pairwise.adonis(phylo_FMT2, subset_meta_FMT2_arrange$type, perm=999)
pairwise%>%
  tt()
```

***NMDS***
```{r beta_neutral_nmds_FMT2, warning=FALSE, comments="", message=FALSE, results="hide"}
 richness_FMT2 %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE, trace = FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(subset_meta_FMT2, by = join_by(sample == Tube_code)) %>%
			group_by(type) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type)) +
        scale_color_manual(name="Type",
                       breaks=c("Cold_control", "Warm_control", "Cold_intervention"),
                       labels=c("Cold_control", "Warm_control", "Cold_intervention"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y = "NMDS2", x="NMDS1 \n Richness beta diversity") +
				theme_classic() +
				theme(legend.position="none")

neutral_FMT2 %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE, trace = FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(subset_meta_FMT2, by = join_by(sample == Tube_code))%>%
			group_by(type) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type)) +
        scale_color_manual(name="Type",
                       breaks=c("Cold_control", "Warm_control", "Cold_intervention"),
                       labels=c("Cold_control", "Warm_control", "Cold_intervention"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y = "NMDS2", x="NMDS1 \n Neutral beta diversity") +
				theme_classic() +
				theme(legend.position="none")

phylo_FMT2 %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE, trace = FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(subset_meta_FMT2, by = join_by(sample == Tube_code))%>%
			group_by(type) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type)) +
				scale_color_manual(name="Type",
                       breaks=c("Cold_control", "Warm_control", "Cold_intervention"),
                       labels=c("Cold_control", "Warm_control", "Cold_intervention"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y= element_blank (), x="NMDS1 \n Phylogenetic beta diversity") +
				theme_classic() +
				theme(legend.position="none")
```

#### Functional differences
***MCI***
```{r gitfs_elements_tm5_FMT2_cold, warning=FALSE, comments="", message=FALSE}
GIFTs_functions_community %>%
  rowMeans() %>%
  as_tibble(., rownames = "sample") %>%
  left_join(sample_metadata, by = join_by(sample == Tube_code)) %>%
  filter(time_point=="FMT2") %>%
  group_by(type) %>%
  summarise(MCI = mean(value), sd = sd(value))
```

```{r gitfs_functional_FMT2_cold, warning=FALSE, comments="", message=FALSE}
MCI_FMT2 <- GIFTs_functions_community %>%
  rowMeans() %>%
  as_tibble(., rownames = "sample") %>%
  left_join(sample_metadata, by = join_by(sample == Tube_code)) %>%
  filter(time_point == "FMT2")

shapiro.test(MCI_FMT2$value)

res.aov <- aov(value ~ type, data = MCI_FMT2)
summary(res.aov)
```

***Functional differences***

CC and CI
```{r comunity_elem_FMT2_cold, comment="", message=FALSE, warning=FALSE}
significant_elements<-element_gift %>%
  filter(type %in% c("Cold_control", "Cold_intervention") & time_point == "FMT2") %>%
  select(-time_point, -type) %>%
  filter(rowSums(across(where(is.numeric), ~ . != 0)) > 0) %>%
  select(Tube_code, where( ~ is.numeric(.) && sum(.) > 0)) %>%
  left_join(sample_metadata[c(1, 6)], by = "Tube_code") %>% 
  pivot_longer(-c(Tube_code, type),
               names_to = "trait",
               values_to = "value") %>%
  group_by(trait) %>%
  summarise(p_value = wilcox.test(value ~ type)$p.value) %>%
  mutate(p_adjust = p.adjust(p_value, method = "BH")) %>%
  filter(p_adjust < 0.05) %>%
  remove_rownames() %>% 
  left_join(., uniqueGIFT_db, by=join_by("trait"=="Code_element"))
```

```{r comunity_elem_fm2_1, comment="", message=FALSE, warning=FALSE}
element_gift_sig <- element_gift %>%
  select(Tube_code, all_of(intersect(
    significant_elements$trait,
    colnames(element_gift)
  ))) %>%
  left_join(., sample_metadata[c(1, 6)], by = join_by(Tube_code == Tube_code)) %>%
  filter(type!="Warm_control")

difference_table <- element_gift_sig %>%
  select(-Tube_code) %>%
  group_by(type) %>%
  summarise(across(everything(), mean)) %>%
  t() %>%
  row_to_names(row_number = 1) %>%
  as.data.frame() %>%
  mutate_if(is.character, as.numeric) %>%
  rownames_to_column(., "Elements") %>%
  left_join(.,uniqueGIFT_db[c(1,3,4)],by = join_by(Elements == Code_element)) %>% 
  arrange(Function) %>% 
  mutate(Difference=Cold_control-Cold_intervention)%>% 
  mutate(group_color = ifelse(Difference <0, "Cold_control","Cold_intervention")) 
```

```{r commun_wilcox_elem_plot_fmt2, comment="", message=FALSE, warning=FALSE, fig.height=6, fig.width=12, fig.fullwidth=TRUE}
difference_table %>%
  ggplot(aes(x=forcats::fct_reorder(Function,Difference), y=Difference, fill=group_color)) + 
  geom_col() +
  scale_fill_manual(values=c("#4477AA","#76b183")) + 
  geom_hline(yintercept=0) + 
  coord_flip()+
  theme(axis.text = element_text(size = 10),
        axis.title = element_text(size = 12),
        legend.position = "right", 
        legend.title = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_line(size = 0.15, linetype = 'solid',
                                        colour = "grey"))+
  xlab("Function") + 
  ylab("Mean difference")
```

CI and WC
```{r comunity_elem_FMT2_warm, comment="", message=FALSE, warning=FALSE}
element_gift %>%
  filter(type %in% c("Warm_control", "Cold_intervention") &
           time_point == "FMT2") %>%
  select(-time_point, -type) %>%
  filter(rowSums(across(where(is.numeric), ~ . != 0)) > 0) %>%
  select(Tube_code, where(~ is.numeric(.) && sum(.) > 0)) %>%
  left_join(sample_metadata[c(1, 6)], by = "Tube_code") %>% 
  pivot_longer(-c(Tube_code,type), names_to = "trait", values_to = "value") %>%
  group_by(trait) %>%
  summarise(p_value = wilcox.test(value ~ type)$p.value) %>%
  mutate(p_adjust=p.adjust(p_value, method="BH")) %>%
  filter(p_adjust < 0.05)%>%
  remove_rownames()
```

CC and WC
```{r comunity_elem_FMT2_control, comment="", message=FALSE, warning=FALSE}
element_gift %>%
  filter(type %in% c("Warm_control", "Cold_control") &
           time_point == "FMT2") %>%
  select(-time_point, -type) %>%
  filter(rowSums(across(where(is.numeric), ~ . != 0)) > 0) %>%
  select(Tube_code, where(~ is.numeric(.) && sum(.) > 0)) %>%
  left_join(sample_metadata[c(1, 6)], by = "Tube_code") %>%
  pivot_longer(-c(Tube_code,type), names_to = "trait", values_to = "value") %>%
  group_by(trait) %>%
  summarise(p_value = wilcox.test(value ~ type)$p.value) %>%
  mutate(p_adjust=p.adjust(p_value, method="BH")) %>%
  filter(p_adjust < 0.05)%>%
  remove_rownames()
```
