# Antimicrobial resistance

```{r load_genome_annotations_amr, warning=FALSE, comments="", message=FALSE}
genome_annotations_amr <- read_tsv("data/gene_annotations_amr.tsv.xz") %>%
    filter(resistance_type=="AMR")

genome_taxonomy_amr <- read_tsv("data/genome_taxonomy_amr.tsv")
```

## Add genome column to the annotations dataframe

```{r genome_annotations_amr, warning=FALSE, comments="", message=FALSE}
genome_annotations_amr <- genome_annotations_amr %>%
  mutate(genome = sub("@.*", "", gene))
```


## Link metadata information to the genomes

```{r genome_annotations_amr_metadata, warning=FALSE, comments="", message=FALSE}
gene_amr_metadata<-genome_annotations_amr %>%
    left_join(merged_data[c(1, 16,23)], by="genome") %>%
    filter(time_point=="Wild") %>%
    left_join(genome_metadata[c(1, 3,6,7,8)], by = "genome")
```

## AMR diversity 

```{r genome_annotations_amr_plot, warning=FALSE, comments="", message=FALSE}
gene_amr_metadata %>%
  count(genome, phylum, resistance_type) %>%
  left_join(merged_data, by="genome") %>%
  ggplot(aes(y = resistance_type, x = species.y, group=species.y, color=species.y, fill=species.y)) +
      geom_boxplot(outlier.shape = NA) +
      geom_jitter(alpha=0.5)


```     
    
gene_amr_metadata %>%
    count(genome, phylum, resistance_type) %>%
    ggplot(aes(x = phylum, y = n, fill = phylum)) +
    geom_boxplot() +
    facet_wrap(~species.y)


gene_amr_metadata %>%
  count(genome, phylum, resistance_type, species.y,resistance_target) %>%
  ggplot(aes(x = phylum, y = n, fill = phylum)) +
  geom_col() +
  facet_wrap(~resistance_target)



  

